Death by Dropdown?<br>Engineer Insightful Shiny Apps<br>with Behavioral Science

### 2025-09-12 conference-materials/2025_posit-conf/draft.qmd
Refactor conference presentation layout and styling...almost done!
- Restructure bidux package description with multi-column layout and centered tagline
- Update code highlighting patterns and add fragment outputs
- Simplify NOTICE section layout and update evidence text
- Reorganize ANTICIPATE section with stacked fragments and improved animations
- Restructure STRUCTURE section with fade transitions and better content organization
- Simplify VALIDATE section with stacked layout
- Consolidate complete workflow examples into streamlined versions
- Split transformation section into before/after

::: {.fragment .center style="color: var(--coral); font-weight: 600; margin-top: 30px;"}
**10x higher conversion with fewer options**
:::

@@

::: {.column width="45%" .fragment .text-xl style="text-align: center; margin-top: 30px;"}
*Your behavioral scientist in the console*
:::
::::

@@

```{r}
#| code-line-numbers: "|3-5|6-10|11-17"
library(bidux)

@@

:::: {.columns}
::: {.column width="45%"}
```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 8

@@

::: {.column width="55%"}
```{r}
#| code-line-numbers: "|2-6|8"
# Minimal input → Maximum insight
notice_stage <- bid_notice(
  previous_stage = interpret_stage,
  problem = "Users are struggling to find key insights",
  evidence = "Users report confusion and telemetry shows low engagement"
)

summary(notice_stage)
```
:::
::::

@@

:::: {.r-stack}
```{r}
anticipate_stage <- bid_anticipate(previous_stage = notice_stage)

# Auto-generates bias mitigations
summary(anticipate_stage)
```

![](img/ui-gain-frame.png){.fragment}

![](img/ui-loss-frame.png){.fragment}

::: {.text-box-overlay-dark .fragment}
**Same Data, Different Story**
:::
::::

@@

```{r}
structure_stage <- bid_structure(previous_stage = anticipate_stage)

# Provides suggestions how how to implement
summary(structure_stage)
```

::: {.fragment}
![](img/f-pattern-heatmap.png){.absolute bottom="50" right="100" width="650"}
:::

@@

```{r}
validate_stage <- bid_validate(previous_stage = structure_stage)

# Generates validation checklist
summary(validate_stage)
```

::: {.fragment}
![](img/ui-summary.png){.absolute bottom="50" right="100" width="650"}
:::

@@

## The Complete Workflow {background-color="#2c3e50"}

::: {.panel-tabset}

### With Telemetry

```{r}
#| code-line-numbers: "|1-2|4-8|10"
# Load and analyze existing usage data
issues <- bid_telemetry("dashboard_telemetry.sqlite")
count(issues, severity, sort = TRUE)

report <- bid_interpret(central_question = "What do users need first?") |>
  bid_notice_issue(issues[1, ]) |> # bridge from telemetry to BID
  bid_anticipate() |>
  bid_structure() |>
  bid_validate() |>
  bid_report(format = "html", include_diagrams = TRUE)
```

### Without Telemetry

```{r}
#| code-line-numbers: "|1-5|6-9|10"
report <- bid_interpret(central_question = "What do users need first?") |>
  bid_notice(
    problem = "Too many filters, unclear defaults",
    evidence = "User feedback and slow engagement"
  ) |>
  bid_anticipate() |>
  bid_structure() |>
  bid_validate() |>
  bid_report(format = "text", include_diagrams = FALSE)
```

@@

![](img/ui-after.png){.fragment}

@@

:::: {.columns .center}
::: {.column width="65%"}
```{r}
#| eval: false

# Install from CRAN
install.packages("bidux")
library(bidux)

# Quick diagnosis without telemetry
result <- bid_interpret("What do users need?") |>
  bid_notice(
    problem = "Death by dropdown",
    evidence = "User complaints"
  ) |>
  bid_anticipate() |>
  bid_structure() |>
  bid_validate()

# View structure recommendations
result$suggestions

# Learn specific concepts (fuzzy matching works!)
bid_concepts("choice overload")

# Get your next steps
result$next_steps
```
:::

@@

:::: {.columns .center}
::: {.column width="25%" .center}
![](img/hex-bidux.png){style="box-shadow: none !important;"}
:::

::: {.column width="50%" .text-display .center}
**Dashboards don't need more features.**<br>
**They need fewer obstacles.**

@@

::: {.column width="25%" .fragment .center}
*Special thanks to posit::conf organizers,<br>
the Shiny community,<br>
and all of you for caring about your users*
:::
::::

>>>>

::: {.fragment .center style="font-weight: 600; margin-top: 30px;"}
**10x higher conversion with fewer options**
:::

@@

::: {.column width="45%"}
:::

::: {.column width="45%" style="margin-top: 150px;"}
:::

::: {.column width="45%" .fragment style="margin-top: 150px;"}
### *Your behavioral scientist in the console* {style="text-align: center !important; color: var(--teal-light) !important;"}
:::
::::

@@

```{r}
#| code-line-numbers: "3|4|5-10|11-18"
#| output-location: fragment

library(bidux)

@@

:::: {.columns}
::: {.column width="40%"}
```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 8

@@

::: {.column width="60%" .fragment}
```{r}
notice_stage <- bid_notice(
  previous_stage = interpret_stage,
  problem = "Users are struggling to find key insights",
  evidence = "Telemetry shows over 75% of filters are rarely used"
)
```
:::
::::

@@

:::: {.r-stack}
::: {.fragment .fade-in-then-out}
```{r}
anticipate_stage <- bid_anticipate(previous_stage = notice_stage)
```
:::

::: {.fragment .fade-in-then-out}

```{r}
# Auto-generates bias mitigations
anticipate_stage$bias_mitigations
```
:::

![](img/ui-gain-frame.png){.fragment auto-animate=true}

![](img/ui-loss-frame.png){.fragment auto-animate=true}

::: {.text-box-overlay-dark .fragment .fade-up}
**Same Data, Different Story**
:::
::::

@@

:::: {.r-stack}
::: {.fragment .fade-in-then-out}
```{r}
structure_stage <- bid_structure(previous_stage = anticipate_stage)
```
:::

::: {.fragment .fade-in}

```{r}
# Provides suggestions how how to implement
str(structure_stage$suggestions[[1]])
```
:::

::: {.fragment .fade-left style="width: 70%"}
![](img/f-pattern-heatmap.png)
:::
::::

@@

:::: {.r-stack}
::: {.fragment .fade-in}
```{r}
validate_stage <- bid_validate(previous_stage = structure_stage)
```
:::

::: {.fragment .fade-up style="width: 70%"}
![](img/ui-summary.png)
:::
::::

@@

## Complete Workflow {background-color="#2c3e50"}

``` {r}
#| echo: false
invisible({
  capture.output({
    issues <- bid_telemetry("dashboard_telemetry.sqlite")
  })
})
```

```{r}
#| eval: false
#| code-line-numbers: "1-2|4-6|7|8-10|11"

# Using telemetry data
issues <- bid_telemetry("dashboard_telemetry.sqlite")

report <- filter(issues, severity == "critical") |>
  slice_head(n = 1) |>
  bid_interpret(central_question = "Which markets are driving performance?") |>
  bid_notice_issue(issues[1, ]) |> # bridge from telemetry to BID
  bid_anticipate() |>
  bid_structure() |>
  bid_validate() |>
  bid_report(format = "markdown", include_diagrams = TRUE)
```

## Complete Workflow {background-color="#2c3e50"}

```{r}
#| code-line-numbers: "1|2-10|11-14|15-17|18|19"
#| output-location: fragment

# Without telemetry data
report <- bid_interpret(
    central_question = "Which markets are driving performance?",
    data_story = list(
      hook = "Q4 revenue hit record high, but satisfaction dipped",
      context = "After aggressive marketing across all regions",
      tension = "West region satisfaction fell 10 points",
      resolution = "Focus retention efforts on underperforming regions"
    )
  ) |>
  bid_notice(
    problem = "Users are struggling to find key insights",
    evidence = "Users report not knowing where to start"
  ) |>
  bid_anticipate() |>
  bid_structure() |>
  bid_validate() |>
  bid_report(format = "text", include_diagrams = FALSE)

print(report)
```

@@

![](img/ui-after.png)

@@

:::: {.columns .center}
::: {.column width="65%"}
```{r}
#| eval: false
#| code-line-numbers: "1-2|4-11|13|15|17"

install.packages("bidux")
library(bidux)

result <- bid_interpret("What do users need?") |>
  bid_notice(
    problem = "Death by dropdown",
    evidence = "User complaints"
  ) |>
  bid_anticipate() |>
  bid_structure() |>
  bid_validate()

summary(result)

bid_concepts("choice overload")

result$next_steps
```
:::

@@

:::: {.columns .center}
::: {.column width="20%" .center}
![](img/hex-bidux.png){style="box-shadow: none !important;"}
:::

::: {.column width="60%" .text-display .center}
**Dashboards don't need more features.**<br>
**They need fewer obstacles.**

@@

::: {.column width="20%" .fragment .center}
*Special thanks to the<br>
posit::conf organizers,<br>
Shiny community,<br>
and all of you for caring about your users*
:::
::::
